---
title: "Fars Functions Vignette"
author: "Gopal Seshadri"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Fars Functions

## fars_read()
  The first function is the **fars_read()** function. This function reads a file, to be precise a  file from Fatality Analysis Reporting System.
  
```{r fars_read}
fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}
```

## make_filename()
  The second function is the **make_filename()** function. This function makes a file name for the given year.
  
```{r make_filename}
make_filename <- function(year) {
  year <- as.integer(year)
  sprintf("accident_%d.csv.bz2", year)
}
```

## fars_read_years()
  The third function is the **fars_read_years()** function. This function takes a vector of years, and for each year in the vector it uses *make_filename()* to make a file name pattern for the given year and *fars_read* to read the file.
  
```{r fars_read_years}
fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = year) %>%
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}
```

## fars_summarize_years()
  The fourth function is the **fars_summarize_years()** function. This function takes a vector of *years* as input. Then reads the list generated by *fars_read_years()* by calling it. Then summarize the data and spreads the data in multiple column.
  
```{r fars_summarize_years}
fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>%
    dplyr::group_by(year, MONTH) %>%
    dplyr::summarize(n = n()) %>%
    tidyr::spread(year, n)
}
```
